From DSL to HPC Component-Based Runtime:\\ A Multi-Stencil DSL Case Study

### 2015-07-29 WOLFHPC15/article/component.tex
WIP Section Components

>>>>

%----------------------------------------
\subsection{Overview}
%----------------------------------------

This section describes both the projection of $\Gamma_{tsp}$ to
components, and the remaining needed components to execute a
multi-stencil application. But, first the section gives an overview of
component models, and then it details \llc, the component model used in
this work.

%----------------------------------------
\subsection{Component Model Overview}
%----------------------------------------
Component model is an domain of software
engineering~\cite{Szyperski:2002:CSB:515228}, which promotes code
re-use and separations of concerns. An application is made by a set of
component; a component being a kind of black box that interacts only
through ports: ports specify the services provided and required.

Component models are known to improved code re-use, separation of
concerns, and maintainability of
applications~\cite{Szyperski:2002:CSB:515228}. More recent works have
also shown that component models can handled performance, scalability,
and portability of applications~\cite{Bernholdt01052006,
  bigot:inria-00388508, l2c}. Many component models exist, each of
them with its own specifications and singularities. In distributed
computing, well known models are CCM~\cite{corba:omg06} (CORBA
Component Model), and GCM~\cite{Baude} (Grid Component Model); in HPC,
there are CCA~\cite{Bernholdt01052006} (Common Component
Architecture), and \llc~\cite{l2c} (Low Level Component), for example.

%----------------------------------------
\subsection{\llc}
%----------------------------------------

The remaining of this paper focuses on \llc, in particular for the
experiments. Let describe it in more details.

\llc is a minimalist \texttt{C++} based HPC component model where a
component extends the concept of class by specifying in its interfaces
the services that it offers ($provide$ ports) and that it needs,
either a single service instance ($use$ ports), or multiple service
instances ($use-multiple$ ports). Services are \texttt{C++}
interfaces. \llc also offer $MPI$ ports that enable components to
share a MPI communicators.
%
As illustrated in Figures~\ref{fig:ports}, a $provide$ port is
represented by a white circle on the left of the component, a $use$
port by a black circle, a $use-multiple$ port by a black circle with a
white $m$ in it, both on the right of the component. MPI PORT!!!

A classical \llc-based application is a static \emph{assembly} of
components made of instances of component and connections between
ports. Such an assembly is describe in LAD (\llc Assembly Descriptor),
an XML dialect~\cite{l2c}.

\begin{figure}[t]
a)\hspace{2mm} \begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
   \node[component] (C) at (0,0) {$C_0$};
   \node[provide] (p) at (-1.5,0) {};
   \node[use] (u) at (1.5,0) {};
   \node[provide,right=1.5cm of u] (p1) {};
   \node[component,right=1.5cm of p1] (C1) {$C_1$};
   \node[use,right=1.5cm of C1] (um) {$m$};

  \path[-]
    (p) edge node {$p$} (C)
    (C.east) edge node {$u$} (u)
    (C1)  edge node {$u'$} (um)
    (p1) edge node {$p'$} (C1);
\end{tikzpicture}
\\\\
b)\hspace{2mm}\begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
   \node[component] (C) at (0,0) {$C_0$};
   \node[provide] (p) at (-1.5,0) {};
   \node[use] (u) at (1.5,0) {};
   \node[component,right=1.5 of u] (C1) {$C_1$};
   \node[use,right=1.5 of C1] (um) {$m$};

  \path[-]
    (p) edge node {} (C)
    (C) edge node {} (u)
    (C1)  edge node {} (um)
    (u) edge node {} (C1);
\end{tikzpicture}
\\\\
c) MPI!!!
\\\\
\caption{a) Component $C_0$ has a provide port ($p$) and a use port ($u$); Component $C_1$ has also a provide port ($p'$) but a use multiple port ($u'$). b) When a use port is connected to a provide port, only the use port is shown and by convention it has to be the left component.}
\label{fig:ports}
\end{figure}

%% In the rest of this paper, when a required service of a \emph{use} (or \emph{use-multiple}) port is filled and linked to a \emph{provide} port in the comonent assembly, only the use port stay visible, as illustrated in Figure~\ref{fig:assembly}. As the use port is on the right of its component, and the provide port on the left, in a component assembly the component on the left uses the provide port of the component on the right.
